---
title: "Week 2 Exercises"
author: "J.Blondeau"
date: "2023-08-31"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Types and Structures in R {.tabset}

In this lesson, we will explore different data types and structures in R and understand how they are used to organize and store data.

### Data Types

#### Numeric

Numeric data types represent numbers, both integers and decimals.

```{r}
# Numeric data example
x <- 10 #integer
y <- 3.14 #floating
```

#### Character
Character data types represent text or strings.

```{r}
# Character data example
species <- "Acropora palmata"
greeting <- "Hello, R!"

```
#### Logical (Boolean)
Logical data types represent binary values: TRUE or FALSE.

```{r}
# Logical data example
is_student <- TRUE
is_bleached <- FALSE

```
### Data Structures

#### Vectors
Vectors are sequences of elements of the same data type.

```{r}
# Vector example
numbers <- c(1, 2, 3, 4, 5)
fruits <- c("apple", "banana", "orange")

```

#### Data Frames
Data frames are tabular structures with rows and columns of different data types.

```{r}
# Data frame example
df <- data.frame(name = c("Alice", "Bob", "Charlie"),
                 age = c(25, 30, 22),
                 is_student = c(TRUE, FALSE, TRUE))

```

#### Lists
Lists are collections of elements, which can be of different data types.

```{r}
# List example
my_list <- list(numbers, fruits, df)

```

### Reading Data from External Files

#### read_csv() Function

The `read_csv()` function from the `readr` package allows you to read data from CSV files.

```{r}
# Reading CSV file using read_csv()
library(readr)

csv_data <- read_csv("data/sttstj_2023_boat_logs.csv", show_col_types = F)

head(csv_data)

```

#### read_xlsx

The `read_xlsx()` function from the `readxl`` package allows you to read data from excel files and specify the sheet in necessary.

```{r}
library(readxl)

xlsx_data <- read_xlsx("data/sttstj_all_sites.xlsx", sheet = "locations")

head(xlsx_data)

```

### Inspecting data

#### class() Function

The `class()` function helps you determine the class (data type) of an R object.

```{r}
# Using class() to determine the class of an object
x <- 5
y <- "Hello, R!"
class(x)   # Output: "numeric"
class(y)   # Output: "character"
```

#### str() Function

The str() function provides a concise summary of an R object's structure.

```{r}
# Using str() to understand the structure of an object
data <- list(name = "Alice", age = 30, is_student = TRUE)
str(data)
```

#### glimpse() Function

The glimpse() function from the dplyr package offers a compact overview of data frames.

```{r}
# Using glimpse() to get an overview of a data frame
library(dplyr)

df <- data.frame(name = c("Alice", "Bob", "Jeremiah"), age = c(25, 30, 46))
glimpse(df)

```

### Exporting Data to External Files

#### write_csv() Function

The `write_csv()` function from the `readr`` package allows you to export data to CSV files.

```{r}
# Creating a data frame
df <- data.frame(name = c("Alice", "Bob", "Jeremiah"),
                 age = c(25, 30,46))

# Exporting data to CSV using write_csv()
library(readr)
write_csv(df, "data/csv_output.csv")
```

#### write_xlsx() Function

The write_xlsx() function from the writr package lets you export data to Excel (.xlsx) files.

```{r}
library(writexl)
# Creating multiple data frames
df <- data.frame(name = c("Alice", "Bob", "Jeremiah"),
                 age = c(25, 30,46))

df_another <- data.frame(name = c("Charlie", "David", "Shirley"),
                          age = c(22, 28, 35))

# Exporting data to Excel using write_xlsx()
write_xlsx(list("Sheet1" = df, "Sheet2" = df_another), "data/excel_output.xlsx")

```

